The role of standing variation in geographic convergent adaptation

### 2015-06-15 standing-variation-paper.tex
removed ugly green boxes from links
minor fix

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage{hyperref}
%\usepackage{latexml} % for \iflatexml
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}  % for special characters in authors' names

@@

Finally, while the form of eqn. \eqref{eqn:defines_chi}, specifying
the characteristic length, is not
particularly intuitive we can use it to ask when we should expect
multiple adaptative alleles within a large geographic region  where our selective pressure
is present (thanks to Sam Yeaman for proposing this
interpretation). Consider the case where this geographic area is a fairly regular
shape of area $G$ and diameter $G^{1/2}$. Denote the total effective population size
over this area by $N = \rho G$, and the standard deviation of dispersal as a
fraction of the diameter of this area by $\sigma_G  = \sigma/
G^{1/2}$. Measuring distance in units of the diameter of $G^{1/2}$,
we expect multiple mutations when $\chi<1$ a
condition which will be met when 
\begin{equation}
N > \frac{\sigma_G}{ \mu \pi \sqrt{2s_b} \left(1 -  \sigma_G \sqrt{2s_b} /\log (1-s_d) \right)},
\end{equation}
this is found by rearranging eqn. \eqref{eqn:defines_chi} having set
$\chi=1$. This nicely shows that $\sigma_G/\mu$ is a primary
determinant of the critical population density necessary for
convergent evolution, when $\sigma_G$ is small. As $\sigma_G
s_b/s_d$ becomes large, e.g. because the allele is not too deleterious
before the environmental shift, we get our standing variation only
case where the dependence of  $\sigma_G$
drops out leaving the critical population size as $1/(2 \mu \pi
s_b/s_d)$

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[authoryear]{natbib}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
%\usepackage{latexml} % for \iflatexml
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage[utf8]{inputenc}  % for special characters in authors' names

@@

Finally, while the form of eqn. \eqref{eqn:defines_chi}, specifying
the characteristic length, is not
particularly intuitive we can use it to ask when we should expect
multiple adaptative alleles within a large geographic region  where our selective pressure
is present (thanks to Sam Yeaman for proposing this
interpretation). Consider the case where this geographic area is a fairly regular
shape of area $G$ and diameter $G^{1/2}$. Denote the total effective population size
over this area by $N = \rho G$, and the standard deviation of dispersal as a
fraction of the diameter of this area by $\sigma_G  = \sigma/
G^{1/2}$. Measuring distance in units of the diameter of $G^{1/2}$,
we expect multiple mutations when $\chi<1$ a
condition which will be met when 
\begin{equation}
N > \frac{\sigma_G}{ \mu \pi \sqrt{2s_b} \left(1 -  \sigma_G \sqrt{2s_b} /\log (1-s_d) \right)},
\end{equation}
this is found by rearranging eqn. \eqref{eqn:defines_chi} having set
$\chi=1$. This nicely shows that $\sigma_G/\mu$ is a primary
determinant of the critical population density necessary for
convergent evolution, when $\sigma_G$ is small. 
As $\sigma_G s_b/s_d$ becomes large, 
e.g., because the allele is not too deleterious before the environmental shift, 
this reduces to the situation with only standing variation,
in which the dependence of  $\sigma_G$
drops out, leaving the critical population size as $1/(2 \mu \pi s_b/s_d)$.
